Extended Data

### 2021-10-13 doc/manuscript/Dow_manuscript_Nature.Rmd
work on structuring discussion

@@

@@

Spring warming at SCBI and Harvard Forest in the study years used here did not result in increased $\Delta DBH$ or summer water stress (SPEI; Supp(?) SPEI Figure). This suggests that even when warmer springs are not associated with greater seasonal water stress, the lack of increased stem growth still prevails at these sites. This could be because the woody growth at these sites are already limited by some mechanism.

@@

Our finding that interannual variation in woody growth is more strongly linked to conditions during the peak growing season than to growing season length aligns with parallel findings for NEE [@zhou_explaining_2016; @fu_maximum_2019].
However, there is also a disconnect between the woody growth responses to spring temperatures observed here and parallel studies on NEE implying that annual C sequestration increases -- at least modestly -- with spring temperatures[@keenan_net_2014] or with the length of the carbon uptake period[@zhou_explaining_2016; @fu_maximum_2019].
We show that the extra C does not go into woody growth, raising the question of where this C goes.

@@

Thus, a distinction between interannual variation and directional change may be critical when considering how directional climate change is likely to affect tree growth and ecosystem C dynamics. 
In addition to the existence of temporal lags between C uptake and release, which introduce a lot of uncertainty[@trumbore_quantifying_2002], trees are likely to  acclimate *to some extent* to warming temperatures[@zhang_drought_2021; @gessler_way_2020], and on even longer time scales, we are likely to see changes in forest composition and adaptation.
If we look across spatial gradients where the latter have had time to play out, we see that longer growing seasons are correlated with greater tree growth[**REF?**], woody productivity[@banburymorgan_global_2021], and NEE [@churkina_spatial_2005].
*Future research will be important to sorting this out.*

Despite remaining uncertainties, our findings contribute substantially to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (Fig. `r DOY_timing`)[@jeong_phenology_2011], an earlier start to net C sequestration[@keenan_net_2014], and, as shown here, earlier initiation of stem growth (Figs. `r schematic_summary`, `r EDF_pheno_Tmax_climwin`).
Yet, an earlier start to the growing season is not consistently associated with a longer peak growing season or increased growth (Figs. `r schematic_summary`, `r tree_rings`), presumably because stem diameter growth is curtailed by C sink limitation[@zani_increased_2020] or drought[@zhang_drought_2021;@dorangeville_drought_2018;@helcoski_growing_2019].

As climate change accelerates and spring temperatures become increasingly warmer, growing seasons will start earlier, but this is unlikely to provide the sustained increase in carbon sequestration that is anticipated in most models (Fig. `r schematic_summary`)[@zani_increased_2020].
Because of this, temperate deciduous forests are likely to sequester less CO~2~ than most models suggest, implying that these models are underestimating future overall atmospheric CO~2~.
(*see [GitHub issue 56 on representation in models](https://github.com/EcoClimLab/growth_phenology/issues/56)*)

>>>>

<!-- Spring warming at SCBI and Harvard Forest in the study years used here did not result in increased $\Delta DBH$ or summer water stress (SPEI; Supp(?) SPEI Figure). This suggests that even when warmer springs are not associated with greater seasonal water stress, the lack of increased stem growth still prevails at these sites. This could be because the woody growth at these sites are already limited by some mechanism. -->

@@

<!-- how our findings fill knowledge gap (1 par, maybe 2)-->

@@

@@

<!-- it remains unclear where extra carbon goes (1 par, maybe 2)-->
Our finding that interannual variation in woody growth is more strongly linked to conditions during the peak growing season than to growing season length aligns with parallel findings for NEE [@zhou_explaining_2016; @fu_maximum_2019].
However, there is also a disconnect between the woody growth responses to spring temperatures observed here and parallel studies on NEE implying that annual C sequestration increases -- at least modestly -- with spring temperatures[@keenan_net_2014] or with the length of the carbon uptake period[@zhou_explaining_2016; @fu_maximum_2019].
We show that the extra C does not go into woody growth, raising the question of where this C goes.

@@

<!-- interannual variation ≠ directional change (1 par)-->
Thus, a distinction between interannual variation and directional change may be critical when considering how directional climate change is likely to affect tree growth and ecosystem C dynamics. 
In addition to the existence of temporal lags between C uptake and release, which introduce a lot of uncertainty[@trumbore_quantifying_2002], trees are likely to  acclimate *to some extent* to warming temperatures[@zhang_drought_2021; @gessler_way_2020], and on even longer time scales, we are likely to see changes in forest composition and adaptation.
If we look across spatial gradients where the latter have had time to play out, we see that longer growing seasons are correlated with greater tree growth[**REF?**], woody productivity[@banburymorgan_global_2021], and NEE [@churkina_spatial_2005].
*Future research will be important to sorting this out.*

<!--"contributions beyond" (1 par)-->
Despite remaining uncertainties, our findings contribute substantially to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (Fig. `r DOY_timing`)[@jeong_phenology_2011], an earlier start to net C sequestration[@keenan_net_2014], and, as shown here, earlier initiation of stem growth (Figs. `r schematic_summary`, `r EDF_pheno_Tmax_climwin`).
Yet, an earlier start to the growing season is not consistently associated with a longer peak growing season or increased growth (Figs. `r schematic_summary`, `r tree_rings`), presumably because stem diameter growth is curtailed by C sink limitation[@zani_increased_2020] or drought[@zhang_drought_2021;@dorangeville_drought_2018;@helcoski_growing_2019].

<!--conclusion (1 par)-->
As climate change accelerates and spring temperatures become increasingly warmer, growing seasons will start earlier, but this is unlikely to provide the sustained increase in carbon sequestration and ensuant negative climate change feedback that is anticipated in most models (Fig. `r schematic_summary`)[@buermann_widespread_2018; @zani_increased_2020;@keenan_net_2014].
Rather, the dominant effect of rising temperatures on forest woody productivity will be a negative effect of high summer temperatures (Fig. `r tree_rings`), which constitutes a positive feedback to climate change.
